---
title: "SDS 291 Final"
author: "Julianna Alvord, Emily Halstead, Sabrina Sayasith"
date: "12/7/2017"
output: pdf_document
---

#This is going to be the working code for our document
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(mosaic)
library(tidyverse)
library(stringr)

MentalHealth <- read_csv("survey.csv")
glimpse(MentalHealth)
```

##Data Cleaning
# Modifying Age
limit age between 5-117
if outside of range, set as NaN (not a number)

```{r}
MentalHealth$Age[MentalHealth$Age < 5 | MentalHealth$Age > 117] <- NaN
```

#State for country outside of the United States
```{r}
#if the country is not the US, set state as Outside
MentalHealth$state[MentalHealth$Country != "United States"] <- "Outside US"
```

#Setting United States Regions and Divisions 
```{r}
#if the state is set as "Outside US", set region as "Outside US"
#if state doesnt match any of the valid states and is not "Outside US", Set as "Invalid State"
Northeast <- c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "PA")
NewEngland <- c("CT", "ME", "MA", "NH", "RI", "VT")
MidAtlantic <- c("NJ", "NY", "PA")
Midwest <- c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD")
EastNorthCentral <- c("IL", "IN", "MI", "OH", "WI")
WestNorthCentral <- c("IA", "KS", "MN", "MO", "NE", "ND", "SD")
South <- c("DE", "FL", "GA", "MD", "NC", "SC", "VA", "DC", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX")
SouthAtlantic <- c("DE", "FL", "GA", "MD", "NC", "SC", "VA", "DC", "WV")
EastSouthCentral <- c("AL", "KY", "MS", "TN")
WestSouthCentral <- c("AR", "LA", "OK", "TX")
West <- c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA")
Mountain <- c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY")
Pacific <- c("AK", "CA", "HI", "OR", "WA")
MentalHealth <- MentalHealth %>% mutate(Region = ifelse(state %in% Northeast, "Northeast", ifelse(state %in% Midwest, "Midwest", ifelse(state %in% South, "South", ifelse(state %in% West, "West", ifelse(state == "Outside US", "Outside US", "Invalid State"))))))

# Set the Region
MentalHealth <- MentalHealth %>% mutate(Division = ifelse(state %in% NewEngland, "New England", ifelse(state %in% MidAtlantic, "Mid Atlantic", ifelse(state %in% EastNorthCentral, "East North Central", ifelse(state %in% WestNorthCentral, "West North Central", ifelse(state %in% SouthAtlantic, "South Atlantic", ifelse(state %in% EastSouthCentral, "East South Central", ifelse(state %in% WestSouthCentral, "West South Central", ifelse(state %in% Mountain, "Mountain", ifelse(state %in% Pacific, "Pacific", ifelse(state == "Outside US", "Outside US", "Invalid State")))))))))))
```

#Gender
```{r}
MentalHealth$Gender <- str_to_upper(MentalHealth$Gender)

MentalHealth$Gender <- str_trim(MentalHealth$Gender)

MentalHealth <- MentalHealth %>%
  mutate(Gender_Grouped = 
           ifelse(Gender %in% c("CIS FEMALE", "FEMAIL", "FEMALE", "FEMALE (TRANS)", "TRANS WOMAN", "WOMAN", "CIS-FEMALE/FEMME", "F", "FEMAKE", "FEMALE (CIS)", "TRANS-FEMALE"), "Female",
            ifelse(Gender %in% c("CIS MAN", "M", "MAILE", "MAL", "MALE (CIS)", "MALE-ISH", "MAN", "CIS MALE", "GUY (-ISH) ^_^", "MAIL", "MAKE", "MALE", "MALE LEANING ANDROGYNOUS", "MALR", "MSLE", "OSTENSIBLY MALE, UNSURE WHAT THAT REALLY MEANS", "SOMETHING KINDA MALE?"), "Male", 
              ifelse(is.na(Gender), "NA",
                ifelse(Gender %in% c("GENDERQUEER", "NON-BINARY", "QUEER/SHE/THEY", "ANDROGYNE", "FLUID", "QUEER"),"Gender Non-Conforming", "Other")))))

tally(MentalHealth$Gender_Grouped)

MentalHealth <- MentalHealth %>%
  mutate(Male = 
              ifelse(Gender_Grouped %in% c("Female", "Gender Non-Conforming"), "Not_Male",
                  ifelse(Gender_Grouped == "Male", "Male",
                  ifelse(Gender_Grouped == "NA", "NA", "Other"))))

MentalHealth <- MentalHealth %>%
  filter(Male %in% c("Not_Male", "Male"))
```

#Treatment to binary 0 or 1
Currently, treatment is coded as 'Yes' or 'No'. We want Yes = 1 and No = 0, so we can use it in glm function.
```{r}
MentalHealth <- MentalHealth %>% mutate(treatmentBinary = ifelse(treatment=="Yes", 1, 0))
```

#Starting Results
```{r}
logm1 <- glm(treatmentBinary ~ Male + Age + Division + leave + tech_company, data=MentalHealth, family=binomial)
summary(logm1)
```


# Exploratory Data Analysis
### Treatment, Age, and Gender Identity
```{r}
library(ggplot2)
```

```{r}
treatmentAgeGenderBox <- ggplot(data = MentalHealth, aes(x= treatment, y = Age, fill = Gender_Grouped)) + geom_boxplot()
treatmentAgeGenderBox
```

# 

```{r}
treatmentAgeBox <- ggplot(data = MentalHealth, aes(x= treatment, y = Age)) + geom_boxplot()
treatmentAgeBox
```

There doesn't seem to be a significant difference in the ages of those who get treatment and those who don't.

### Is there a difference between regions? region, treatment, gender within region?
```{r}
divAgeBox <- ggplot(data = MentalHealth, aes(x = Region, y = Age, colour = treatment)) + geom_boxplot()
divAgeBox
```

```{r}
boxNo_Employees <- ggplot(data = MentalHealth, aes(x = no_employees, y = Age, fill = treatment)) + geom_boxplot()
boxNo_Employees
```


```{r}
boxLeave <- ggplot(data = MentalHealth, aes(x = leave, y = Age, fill = treatment)) + geom_boxplot()
boxLeave
```


## Chi-Square. Do the distributions of categorical variables differ from each other?

```{r}
library(reshape2)
#Create contingency tables for chi square
treatDivision <- dcast(MentalHealth, treatment~Division, fill = 0) #treatment vs Division
treatRegion <- dcast(MentalHealth, treatment~Region, fill = 0) #treatment vs Region
treatNo_Employees <- dcast(MentalHealth, treatment~no_employees, fill = 0) #treatment vs No_Employees
treatLeave <- dcast(MentalHealth, treatment~leave, fill = 0) #treatment vs Leave
```
 
### Chi-Square and Odds Ratio:  treatment vs leave

```{r}
#chi square test between treatment and Leave
#For treatment vs no treatment, is there a significant difference in the distributions of the 4 different "leave" groups (Don't Know, Somewhat difficult, Somewhat easy, Very difficult),
#Is "leave"" related to treatment? Is the distribution of leave between treatment and no treatment what we would expect from chance?
#H0:The variables are independent. 
#HA: The variables are not independent. 
chisq.test(treatLeave[c(2:5)])
#p-value = 1.941e-06, smaller than our .05 confidence level
#We reject the null hypothesis that the variables are independent. There is a relationship between treatment and leave.

#melt dataframe for plotting
treatLeaveMelt <- melt(treatLeave, id.vars = 'treatment')
#visualization of the differences
ggplot(treatLeaveMelt, aes(treatment, value)) + geom_bar(aes(fill=variable), position = "dodge", stat="identity")
#all in all, the distributions for yes treatment vs no treatment look pretty similar

#odds the person has sought treatment for a mental illness depending on how easy it is for them to leave for a mental health condition
Odds_DontKnowLeave <- 232/309
Odds_SomewhatDifficultLeave <- 82/44
Odds_SomewhatEasyLeave <- 130/135
Odds_VeryDifficultLeave <- 65/31 
Odds_VeryEasyLeave <- 101/103

#Odds Ratio Comparisons
OR_VDifficult_VEasy <- Odds_VeryDifficultLeave/Odds_VeryEasyLeave
OR_VDifficult_VEasy
#OR= 2.138294. The odds of seeking treatment were 2.138 times higher with those who said that it is very diffiicult for them to leave for a mental health condition than with those who said it would be very easy to leave
OR_VEasy_SomeEasy <- Odds_VeryEasyLeave/Odds_SomewhatEasyLeave
OR_VEasy_SomeEasy
#odds are pretty similar between Very easy and Somewhat easy groups
#OR = 1.018, The odds of seeking treatment were 1.02 times higher with those who said it would very easy for them leave than with those who said it would be somewhat easy
OR_SomeDifficult_SomeEasy <- Odds_SomewhatDifficultLeave/Odds_SomewhatEasyLeave
OR_SomeDifficult_SomeEasy
#OR = 1.93, The odds of seeking treatment were 1.93 times higher with those who said it would be somewhat difficult for them to leave than with those who said it would be somewhat easy
```

### Chi-Square and Odds Ratio: treatment vs No_Employees

```{r}
#chi square test between treatment and no_employees
#For treatment vs no treatment, is there a significant difference in the distributions of the 6 different no_employee groups (1-5, 6-25, 26-100, 100-500, 500-1000, more than 1000)
#Is no_employees related to treatment? Is the distribution of no-employees between treatment and no treatment what we would expect from chance?
#H0:The variables are independent. 
#HA: The variables are not independent. 
chisq.test(treatNo_Employees[c(2:7)])
#p-value = .0165, Greater than the .05 confidence level
#We cannot reject the null hypothesis that the variables are independent
```

### Chi-Square and Odds Ratio: treatment vs. Region
```{r}
#chi square test between treatment and Region
#For treatment vs no treatment, is there a significant difference in the distributions of the 6 different region groups (Midwest, Northeast, OutsideUS, South, West, NA)
#Is no_employees related to treatment? Is the distribution of no-employees between treatment and no treatment what we would expect from chance?
#H0:The variables are independent. 
#HA: The variables are not independent. 
chisq.test(treatRegion[c(2:7)])
#p-value = .0165, Greater than the .05 confidence level
#We cannot reject the null hypothesis that the variables are independent
```

### Step-wise
```{r}
#fullMSE <- (summary(logm1)$sigma)^2
#step(null_mod, scope = list(upper = logm1),
#scale = fullMSE, direction = "both")
#where logm1 is the full model and null_mod was treatmentBinary~1

#remove NA values from data set
#comments variable has a lot of NA values, it's probably OK to remove as the answers are so variable/they are opinions. we are not doing nlp.
MentalHealth_NoNAs <- MentalHealth[c(-27)]
MentalHealth_NoNAs <- MentalHealth_NoNAs[complete.cases(MentalHealth_NoNAs), ] #left with 960 observations
#null model
model.null <- glm(treatmentBinary~1, data = MentalHealth_NoNAs, family = binomial(link="logit"))
#full model
model.full = glm(treatmentBinary~Age + Gender_Grouped + Division + Region + obs_consequence + mental_vs_physical + phys_health_interview + mental_health_interview + supervisor + coworkers + phys_health_consequence + mental_health_consequence + leave + anonymity + seek_help + wellness_program + care_options + benefits + tech_company + remote_work + no_employees + work_interfere + family_history + self_employed + state , data = MentalHealth_NoNAs, family = binomial(link="logit"), control = list(maxit = 50)) # surprised this works with non numeric variables
#was getting: glm.fit: algorithm did not convergeglm.fit: fitted probabilities numerically 0 or 1 occurred
#removed country and it doesn't give error now....
step(model.null,
     scope = list(upper=model.full),
             direction="both",
             test="Chisq",
             data=MentalHealth_NoNAs)
```

### Classification Tree

```{r}
library(rpart)
library(rpart.plot)
#source: https://www.statmethods.net/advstats/cart.html
decisionTree <- rpart(treatmentBinary~Age + Gender_Grouped + Division + Region + obs_consequence + mental_vs_physical + phys_health_interview + mental_health_interview + supervisor + coworkers + phys_health_consequence + mental_health_consequence + leave + anonymity + seek_help + wellness_program + care_options + benefits + tech_company + remote_work + no_employees + work_interfere + family_history + self_employed + state + Country , data = MentalHealth_NoNAs, method = "class", control=rpart.control(minsplit=20, cp=0.001))
#may be overfitted
decisionTree <- prune(decisionTree, cp = .01)
plotTree <- rpart.plot(decisionTree)
#work_interfere seems to be the most important in predicting that someone won't get treatment. if someone says that work never inteferes, most likely to not get treatment

```

