---
title: "SDS 291 Final"
author: "Julianna Alvord, Emily Halstead, Sabrina Sayasith"
date: "12/7/2017"
output: pdf_document
---

#This is going to be the working code for our document
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(mosaic)
library(tidyverse)
library(stringr)

MentalHealth <- read_csv("survey.csv")
glimpse(MentalHealth)
```

##Data Cleaning
# Modifying Age
limit age between 5-117
if outside of range, set as NaN (not a number)

```{r}
MentalHealth$Age[MentalHealth$Age < 5 | MentalHealth$Age > 117] <- NaN
```

#State for country outside of the United States
```{r}
#if the country is not the US, set state as Outside
MentalHealth$state[MentalHealth$Country != "United States"] <- "Outside US"
```

#Setting United States Regions and Divisions 
```{r}
#if the state is set as "Outside US", set region as "Outside US"
#if state doesnt match any of the valid states and is not "Outside US", Set as "Invalid State"
Northeast <- c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "PA")
NewEngland <- c("CT", "ME", "MA", "NH", "RI", "VT")
MidAtlantic <- c("NJ", "NY", "PA")
Midwest <- c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD")
EastNorthCentral <- c("IL", "IN", "MI", "OH", "WI")
WestNorthCentral <- c("IA", "KS", "MN", "MO", "NE", "ND", "SD")
South <- c("DE", "FL", "GA", "MD", "NC", "SC", "VA", "DC", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX")
SouthAtlantic <- c("DE", "FL", "GA", "MD", "NC", "SC", "VA", "DC", "WV")
EastSouthCentral <- c("AL", "KY", "MS", "TN")
WestSouthCentral <- c("AR", "LA", "OK", "TX")
West <- c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA")
Mountain <- c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY")
Pacific <- c("AK", "CA", "HI", "OR", "WA")
MentalHealth <- MentalHealth %>% mutate(Region = ifelse(state %in% Northeast, "Northeast", ifelse(state %in% Midwest, "Midwest", ifelse(state %in% South, "South", ifelse(state %in% West, "West", ifelse(state == "Outside US", "Outside US", "Invalid State"))))))

# Set the Region
MentalHealth <- MentalHealth %>% mutate(Division = ifelse(state %in% NewEngland, "New England", ifelse(state %in% MidAtlantic, "Mid Atlantic", ifelse(state %in% EastNorthCentral, "East North Central", ifelse(state %in% WestNorthCentral, "West North Central", ifelse(state %in% SouthAtlantic, "South Atlantic", ifelse(state %in% EastSouthCentral, "East South Central", ifelse(state %in% WestSouthCentral, "West South Central", ifelse(state %in% Mountain, "Mountain", ifelse(state %in% Pacific, "Pacific", ifelse(state == "Outside US", "Outside US", "Invalid State")))))))))))
```

#Gender
```{r}
MentalHealth$Gender <- str_to_upper(MentalHealth$Gender)

MentalHealth$Gender <- str_trim(MentalHealth$Gender)

MentalHealth <- MentalHealth %>%
  mutate(Gender_Grouped = 
           ifelse(Gender %in% c("CIS FEMALE", "FEMAIL", "FEMALE", "FEMALE (TRANS)", "TRANS WOMAN", "WOMAN", "CIS-FEMALE/FEMME", "F", "FEMAKE", "FEMALE (CIS)", "TRANS-FEMALE"), "Female",
            ifelse(Gender %in% c("CIS MAN", "M", "MAILE", "MAL", "MALE (CIS)", "MALE-ISH", "MAN", "CIS MALE", "GUY (-ISH) ^_^", "MAIL", "MAKE", "MALE", "MALE LEANING ANDROGYNOUS", "MALR", "MSLE", "OSTENSIBLY MALE, UNSURE WHAT THAT REALLY MEANS", "SOMETHING KINDA MALE?"), "Male", 
              ifelse(is.na(Gender), "NA",
                ifelse(Gender %in% c("GENDERQUEER", "NON-BINARY", "QUEER/SHE/THEY", "ANDROGYNE", "FLUID", "QUEER"),"Gender Non-Conforming", "Other")))))

tally(MentalHealth$Gender_Grouped)

MentalHealth <- MentalHealth %>%
  mutate(Male = 
              ifelse(Gender_Grouped %in% c("Female", "Gender Non-Conforming"), "Not_Male",
                  ifelse(Gender_Grouped == "Male", "Male",
                  ifelse(Gender_Grouped == "NA", "NA", "Other"))))

MentalHealth <- MentalHealth %>%
  filter(Male %in% c("Not_Male", "Male"))
```

#Treatment to binary 0 or 1
Currently, treatment is coded as 'Yes' or 'No'. We want Yes = 1 and No = 0, so we can use it in glm function.
```{r}
MentalHealth2 <- MentalHealth %>% mutate(treatmentBinary = ifelse(treatment=="Yes", 1, 0))
```

#Starting Results
```{r}
tally(MentalHealth$tech_company)

logm1 <- glm(treatment ~ Male + Age + Division + leave + tech_company, data=MentalHealth, family=binomial)
summary(logm1)
```
